Necessity is the mother of genetic invention

### 2017-11-03 ms.tex
discussion

\subsection*{Mutation modification model 1.}\label{mut_model_1}

@@

\subsection*{Mutation modification model 2.}\label{mut_model_2}

@@

\subsection*{Migration modification model.}\label{mig_model_2}

@@

We turn our attention to a modifier that concurrently controls the transition out of a subset of types, $K$.
For example, \citet[Appendix~B]{Ram2012} considered individuals that are grouped by the number of their accumulated deleterious mutations, $k$, and a modifier that regulates the mutation rates in individuals with at least $\pi$ deleterious mutations.
Similarly, if the transition rates are controlled by a threshold rule 
\begin{equation}
C_k = \begin{cases}
\tau, & k \in K \\
1, & k \not\in K
\end{cases},
\end{equation}
where $\tau > 0$, then the sensitivity of the stable population mean fitness to changes in the
rate of transition from types $K$ is
\begin{equation}
\begin{aligned}
\frac{\partial \bar{\omega}^*}{\partial \tau} = 
u\tr \Big( \sum_{k \in \mathbb{K{}}}{\frac{\partial (I - C + MC)D}{\partial \tau}} \Big) v = \\
\frac{1}{\tau} \sum_{k \in \mathbb{K}}{u_k v_k (\bar{\omega}^* - D_k)},
\end{aligned}
\end{equation}
which can be simplified to
\begin{multline}
sign \frac{\partial \bar{\omega}^*}{\partial \tau} = 
sign \sum_{k \in \mathbb{K}}{u_k v_k (\bar{\omega}^* - D_k)} = \\ 
sign \Big(\bar{\omega}^* - \frac{\sum_{k \in \mathbb{K}}{u_k v_k D_k}}{\sum_{k \in \mathbb{K}}{u_k v_k}}\Big), 
\end{multline}
a generalization of Eq.~\ref{eq:sign_theorem}.
Therefore, increased transition from types $K$ will increase the stable population mean fitness if and only if the average fitness of individuals descended from types $K$ is below the stable population mean fitness.

% Discussion
\section*{Discussion}
\textbf{TODO}: Departure from the \emph{reduction principle}

Eq.~\ref{eq:theorem} is a generalization of a result from \citet[Eq.~4]{Ram2012}.
Translated to the notation presented here, \citeauthor{Ram2012} modeled the accumulation of deleterious mutations in a population at the mutation-selection balance, assuming no more than one mutation occurs per individual per generation... \textbf{TODO} continue this...

Using Eq.~\ref{eq:Caswells_formula} in \nameref{sec:AppB} and a recursion on the ratios of the reproductive values \cite[see][eqs.~A5-6]{Ram2012}, they concluded that at the mutation-selection balance, if individuals with below-average fitness ($D_k < \bar{\omega}^*$) increase their mutation rate, then the population mean fitness will increase.

@@

\textbf{TODO}

@@

\bibliographystyle{plainnat}
\bibliography{ms}

>>>>

\subsection*{Mutation modification model 1.}\label{sec:mut_model_1}

@@

\subsection*{Mutation modification model 2.}\label{sec:mut_model_2}

@@

\subsection*{Migration modification model.}\label{sec:mig_model_1}

@@

We turn our attention to a modifier that concurrently controls the transition out of a subset of types, $K$.
For example, \citet[Appendix~B]{Ram2012} considered individuals that are grouped by the number of their accumulated deleterious mutations, $k$, and a modifier that regulates the mutation rates in individuals with at least $\pi$ deleterious mutations.
Similarly, if the transition rates are controlled by a threshold rule 
\begin{equation}
C_k = \begin{cases}
\tau, & k \in K \\
1, & k \not\in K
\end{cases},
\end{equation}
where $\tau > 0$, then the sensitivity of the stable population mean fitness to changes in the
rate of transition from types $K$ is
\begin{equation}
\begin{aligned}
\frac{\partial \bar{\omega}^*}{\partial \tau} = 
u\tr \Big( \sum_{k \in \mathbb{K{}}}{\frac{\partial (I - C + MC)D}{\partial \tau}} \Big) v = \\
\frac{1}{\tau} \sum_{k \in \mathbb{K}}{u_k v_k (\bar{\omega}^* - D_k)},
\end{aligned}
\end{equation}
which can be simplified to
\begin{multline}\label{eq:sign_theorem_threshold}
sign \frac{\partial \bar{\omega}^*}{\partial \tau} = 
sign \sum_{k \in \mathbb{K}}{u_k v_k (\bar{\omega}^* - D_k)} = \\ 
sign \Big(\bar{\omega}^* - \frac{\sum_{k \in \mathbb{K}}{u_k v_k D_k}}{\sum_{k \in \mathbb{K}}{u_k v_k}}\Big), 
\end{multline}
a generalization of Eq.~\ref{eq:sign_theorem}.
Therefore, increased transition from types $K$ will increase the stable population mean fitness if and only if the average fitness of individuals descended from types $K$ is below the stable population mean fitness.

% Discussion
\section*{Discussion}

We have shown that under constant-viability selection and in an effectively infinite haploid populations at an equilibrium, the stable population mean fitness increases if below-fitness individuals increase the rate at which variation is generated.
We have given several examples on variation-generating processes for which this principle applies -- mainly mutation and migration -- but our model is general enough to possibly apply to other processes as well.
For example, the \emph{reduction principle} appears some ecological models of dispersal, and~\citet{Gueijman2013} have demonstrated that even in homogeneous environments, fitness-associated dispersal increases the population mean fitness and is favored by selection over uniform dispersal.
Similarly, if the transmission fidelity of culturally-transmitted traits depends on the type of the type or fitness of the transmitting individual, we expect that our \emph{mean fitness principle} will hold.

Eq.~\ref{eq:theorem} is a generalization of a result from \citet[Eq.~4]{Ram2012}.
\citeauthor{Ram2012} modeled the accumulation of mutant alleles in a population evolving at the mutation-selection balance (see \nameref{sec:mut_model_2}).
Using Eq.~\ref{eq:Caswells_formula} in \nameref{sec:AppB} and a recursion on the ratios of the reproductive values \cite[see][eqs.~A5-6]{Ram2012}, they concluded that at the mutation-selection balance, if individuals with below-average fitness ($D_k < \bar{\omega}^*$) increase their mutation rate, then the population mean fitness will increase -- a result generalized by our \emph{mean fitness principle} in Eqs.~\ref{eq:sign_theorem} and~\ref{eq:sign_theorem_threshold}.

Our analysis focuses on populations at equilibrium.
Nevertheless, it has been demonstrated that a modifier which increases the mutation rate of maladapted individuals can be favored by selection~\citep{Ram2012} and reduce the adaptation rate~\citep{Ram2014}.
Similar results have been demonstrated for a modifier that increases the recombination rate in maladapted individuals~\citep{Hadany2003a,Hadany2003b}.
Importantly, there are empirical evidence both for stress-induced mutagenesis~\citep{Rosenberg2012} and for condition-dependent recombination and sex~\citep{Ram2016}.

@@

General examples of departures from the \emph{reduction principle} are rare~\citep{Altenberg2017}.
However, it has been demonstrated that higher mutation, recombination, and migration rates are favored by selection in populations evolving under fluctuating selection; see~\citet{Carja2014} and references therein.
Here we have provided another general example, which suggests that if maladapted or below-average individuals increase the rate at which variation is generated, the population mean fitness will increase.

@@

\bibliographystyle{plainnat}
%\bibliography{/Users/yoavram/Documents/library}
\bibliography{ms}
